Quality measure notes

### 2021-08-28 StateOfPractice/Ao-Report/11Introduction.tex
MI report - pointed answers to research questions

\begin{description}
\item[RQ1.] What artifacts are present in current software projects? What role does documentation play in the projects? What are the developers' attitude toward it?
\item[RQ2.] What tools are used in the development of current software packages?
\item[RQ3.] What principles, processes, and methodologies are used in the development of current software packages?
\item[RQ4.] What are the pain points for developers working on research software projects? What aspects of the existing processes, methodologies, and tools do they consider as potentially needing improvement? What changes to processes, methodologies, and tools can improve software development and software quality?
\item[RQ5.] What is the current status of the following software qualities for the projects? What actions have the developers taken to address them?
\begin{itemize}
	\item Installability
	\item Correctness \& Verifiability
	\item Reliability
	\item Robustness
	\item Usability
	\item Maintainability
	\item Reusability
	\item Understandability
	\item Visibility/Transparency
	\item Reproducibility
\end{itemize}
\item[RQ6.] How does software designated as high quality by this methodology compare with top-rated software by the community?
\end{description}

>>>>

\begin{description}
\item[RQ1.] \hypertarget{rq1}What artifacts are present in current software projects? What role does documentation play in the projects? What are the developers' attitude toward it?
\item[RQ2.] \hypertarget{rq2}What tools are used in the development of current software packages?
\item[RQ3.] \hypertarget{rq3}What principles, processes, and methodologies are used in the development of current software packages?
\item[RQ4.] \hypertarget{rq4}What are the pain points for developers working on research software projects? What aspects of the existing processes, methodologies, and tools do they consider as potentially needing improvement? What changes to processes, methodologies, and tools can improve software development and software quality?
\item[RQ5.] \hypertarget{rq5}What is the current status of the following software qualities for the projects? What actions have the developers taken to address them?
\begin{itemize}
	\item Installability
	\item Correctness \& Verifiability
	\item Reliability
	\item Robustness
	\item Usability
	\item Maintainability
	\item Reusability
	\item Understandability
	\item Visibility/Transparency
	\item Reproducibility
\end{itemize}
\item[RQ6.] \hypertarget{rq6}How does software designated as high quality by this methodology compare with top-rated software by the community?
\end{description}

### 2021-08-28 StateOfPractice/Ao-Report/14MeasurementResults.tex
MI report - pointed answers to research questions

As discussed in Section \ref{sec_applying_method}, we applied our method to the MI domain. This section shows the summary of the measurement results. The detailed data can be found in the repository \hyperlink{https://data.mendeley.com/datasets/k3pcdvdzj2/1}{https://data.mendeley.com/datasets/k3pcdvdzj2/1}.

@@

@@

Table \ref{tab_user_support_model} shows the user support models by the number of projects using them. Maybe not every team intended to use GitHub issues to answer users' questions, but many users use them to seek help.

@@

We marked \textit{3D Slicer} with a much higher score than others because it did very well at closing the identified issues, and more importantly, we found it to have the most comprehensive artifacts. For example, as far as we could find out, only a few of the 29 projects had a project plan, developer's manual, or API documentation, and only \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji} included all three documents. Meanwhile, \textit{3D Slicer} has a much higher percentage of closed issues (91.65\%) than \textit{ImageJ} (52.49\%) and \textit{Fiji} (63.79\%). Table \ref{tab_maintainability_docs} shows which projects had these documents, in the descending order of their \textit{maintainability} scores.

@@

27 of the 29 projects used git as the version control tool; \textit{AMIDE} used Mercurial; \textit{Gwyddion} used Subversion. 24 projects used GitHub for their repositories; \textit{XMedCon
}, \textit{AMIDE}, and \textit{Gwyddion} used SourceForge; \textit{DicomBrowser} and \textit{3DimViewer} used BitBucket.

@@

Table \ref{tab_Visibility/Transparency_docs} shows the projects which had documents for the development process, project status, development environment, and release notes, in the descending order of their \textit{visibility/transparency} scores.

>>>>

As discussed in Section \ref{sec_applying_method}, we applied our method to the MI domain. This section shows the summary of the measurement results. The detailed data can be found in the repository \hyperlink{https://data.mendeley.com/datasets/k3pcdvdzj2/1}{https://data.mendeley.com/datasets/k3pcdvdzj2/1}. This section contains part of the answers to \hyperlink{rq5}{RQ5}.

@@

In this section, the information about CI/CD tools is part of the answers to \hyperlink{rq2}{RQ2}, and the information about software testing and documentation is part of the answers to \hyperlink{rq3}{RQ3}.

@@

Table \ref{tab_user_support_model} shows the user support models by the number of projects using them. This table contains part of the answers to \hyperlink{rq2}{RQ2}. Maybe not every team intended to use GitHub issues to answer users' questions, but many users use them to seek help.

@@

We marked \textit{3D Slicer} with a much higher score than others because it did very well at closing the identified issues, and more importantly, we found it to have the most comprehensive artifacts. For example, as far as we could find out, only a few of the 29 projects had a project plan, developer's manual, or API documentation, and only \textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji} included all three documents. Meanwhile, \textit{3D Slicer} has a much higher percentage of closed issues (91.65\%) than \textit{ImageJ} (52.49\%) and \textit{Fiji} (63.79\%). Table \ref{tab_maintainability_docs} shows which projects had these documents, in the descending order of their \textit{maintainability} scores. This table contains part of the answers to \hyperlink{rq1}{RQ1} and \hyperlink{rq3}{RQ3}.

@@

27 of the 29 projects used git as the version control tool; \textit{AMIDE} used Mercurial; \textit{Gwyddion} used Subversion. 24 projects used GitHub for their repositories; \textit{XMedCon
}, \textit{AMIDE}, and \textit{Gwyddion} used SourceForge; \textit{DicomBrowser} and \textit{3DimViewer} used BitBucket. The information about version control tools is part of the answers to \hyperlink{rq2}{RQ2}.

@@

Table \ref{tab_Visibility/Transparency_docs} shows the projects which had documents for the development process, project status, development environment, and release notes, in the descending order of their \textit{visibility/transparency} scores. This table contains part of the answers to \hyperlink{rq1}{RQ1} and \hyperlink{rq3}{RQ3}.

### 2021-08-28 StateOfPractice/Ao-Report/15InterviewsWithDevelopers.tex
MI report - updated ch 6 and more
MI report - pointed answers to research questions

By asking questions \hyperlink{q9}{9}, \hyperlink{q10}{10}, and \hyperlink{q12}{12}, we tried to identify the pain points during the development process in the eight projects. The pain points include current and past obstacles. We also asked the interviewees how they would solve the problems. Questions 9, 10, and 12 are as follows:

@@

\subsection{Testing Pain Point}
\label{sec_pain_points_3}
The pain point in the \textit{testing} category is \textbf{the lack of access to real-world datasets for testing.}

@@

\section{Documents in the Projects}
\label{sec_interview_documents}
We tried to understand the interviewees' opinions on documentation and the quality of documentations with questions 11 and 19:

@@

Table \ref{tab_doc_tools} lists some of the documentation tools and methods mentioned by the interviewees.

@@

Although some team may have a documented process for accepting new contributions, no one talked about it during the interview. However, most of them mentioned using GitHub and pull requests to manage contributions from the community. The interviewees generally gave very positive feedback on using GitHub. Some also said they had handled the project repository with some other tools, and eventually transferred to git and GitHub. Table \ref{tab_method_receive_contributions} shows the number of times the interviewees mentioned the methods of receiving contributions.

@@

Some interviewees mentioned the project management tools they used, which are in Table \ref{tab_pm_tools}. Generally speaking, they talked about two types:

@@

\section{Discussions on Software Qualities}
\label{sec_interview_software_qualities}
Questions 15--18, and 20 are about the software qualities of \textit{correctness}, \textit{maintainability}, \textit{understandability}, \textit{usability}, and \textit{reproducibility}, respectively. We asked these questions to understand the threats to these qualities and the developers' strategies to improve them. We discuss each quality in a separate section below.

@@

@@

\subsection{Understandability}

\begin{description}
\item[Q17.] \hypertarget{q17}Provide instances where users have misunderstood the software. What, if any, actions were taken to address understandability issues?
\end{description}

@@

Table \ref{tab_q18_strategies_usability} shows the strategies to ensure \textit{usability} by the numbers of interviewees with the answers.

>>>>

By asking questions \hyperlink{q9}{9}, \hyperlink{q10}{10}, and \hyperlink{q12}{12}, we tried to identify the pain points during the development process in the eight projects. The pain points include current and past obstacles. We also asked the interviewees how they would solve the problems. This section contains the answers to \hyperlink{rq4}{RQ4}. Questions 9, 10, and 12 are as follows:

@@

\subsection{Testing Pain Point}
\label{sec_pain_points_3}
The pain point in the \textit{testing} category is \textbf{the lack of access to real-world datasets for testing.} The information about software testing in this section is part of the answers to \hyperlink{rq3}{RQ3}.

@@

\section{Documents in the Projects}
\label{sec_interview_documents}
We tried to understand the interviewees' opinions on documentation and the quality of documentations with questions 11 and 19. The information about documentation is part of the answers to \hyperlink{rq3}{RQ3}.

@@

Table \ref{tab_doc_tools} lists some of the documentation tools and methods mentioned by the interviewees. This table contains part of the answers to \hyperlink{rq2}{RQ2}.

@@

Although some team may have a documented process for accepting new contributions, no one talked about it during the interview. However, most of them mentioned using GitHub and pull requests to manage contributions from the community. The interviewees generally gave very positive feedback on using GitHub. Some also said they had handled the project repository with some other tools, and eventually transferred to git and GitHub. Table \ref{tab_method_receive_contributions} shows the number of times the interviewees mentioned the methods of receiving contributions. This table contains part of the answers to \hyperlink{rq2}{RQ2}.

@@

Some interviewees mentioned the project management tools they used, which are in Table \ref{tab_pm_tools}. This table contains part of the answers to \hyperlink{rq2}{RQ2}. Generally speaking, the interviewees talked about two types:

@@

The information about contribution management and project management in this section is part of the answers to \hyperlink{rq3}{RQ3}.

\section{Discussions on Software Qualities}
\label{sec_interview_software_qualities}
Questions 15--18, and 20 are about the software qualities of \textit{correctness}, \textit{maintainability}, \textit{understandability}, \textit{usability}, and \textit{reproducibility}, respectively. We asked these questions to understand the threats to these qualities and the developers' strategies to improve them. We discuss each quality in a separate section below. This section contains part of the answers to \hyperlink{rq5}{RQ5}.

@@

The information about software testing in this section is part of the answers to \hyperlink{rq3}{RQ3}.

@@

\subsection{Understandability}
\label{sec_interview_understandability}
\begin{description}
\item[Q17.] \hypertarget{q17}Provide instances where users have misunderstood the software. What, if any, actions were taken to address understandability issues?
\end{description}

@@

Table \ref{tab_q18_strategies_usability} shows the strategies to ensure \textit{usability} by the numbers of interviewees with the answers. The information about software testing in this table is part of the answers to \hyperlink{rq3}{RQ3}.

### 2021-08-28 StateOfPractice/Ao-Report/16AnswersToResearchQuestions.tex
MI report - pointed answers to research questions

As mentioned in Section \ref{sec_grading_template}, we search for all the artifacts in a project when measuring \textit{maintainability}. The detailed records of the existing artifacts in the 29 MI projects are at \hyperlink{https://data.mendeley.com/datasets/k3pcdvdzj2/1}{https://data.mendeley.com/datasets/k3pcdvdzj2/1}. Table \ref{tab_artifacts_frequency} summarizes the frequency of the artifacts.

@@

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Section or table & Description \\ \hline
Section \ref{fg_correctness_erifiability_scores}, \ref{sec_pain_points_3}, \ref{sec_interview_correctness}, \ref{sec_interview_usability}, and \ref{sec_interview_reproducibility}; Table \ref{tab_q15_threats_correctness} and \ref{tab_q15_strategies_correctness}. & Software testing \\
Section \ref{fg_correctness_erifiability_scores} and \ref{sec_interview_documents}; Table \ref{tab_maintainability_docs}, \ref{tab_Visibility/Transparency_docs}, and \ref{tab_artifacts_frequency} & Documentation \\
Section \ref{sec_contribution_pm} & Contribution management \\
Section \ref{sec_contribution_pm} & Project management \\ \hline
\end{tabular}
\caption{\label{tab_reference_rq3}Sections and tables with answers to RQ3}
\end{table}

>>>>

As mentioned in Section \ref{sec_grading_template}, we search for all the artifacts in a project when measuring \textit{maintainability}. The detailed records of the existing artifacts in the 29 MI projects are at \hyperlink{https://data.mendeley.com/datasets/k3pcdvdzj2/1}{https://data.mendeley.com/datasets/k3pcdvdzj2/1}. Table \ref{tab_artifacts_frequency} summarizes the frequency of the artifacts. This table also contains part of the answers to \hyperlink{rq3}{RQ3}.

@@

\begin{table}[H]
\centering
\begin{tabular}{ll}
\hline
Section or table & Description \\ \hline
Section \ref{fg_correctness_erifiability_scores}, \ref{sec_pain_points_3}, \ref{sec_interview_correctness}, and \ref{sec_interview_reproducibility}; Table \ref{tab_q18_strategies_usability} & Software testing \\
Section \ref{fg_correctness_erifiability_scores} and \ref{sec_interview_documents}; Table \ref{tab_maintainability_docs}, \ref{tab_Visibility/Transparency_docs}, and \ref{tab_artifacts_frequency} & Documentation \\
Section \ref{sec_contribution_pm} & Contribution management \\
Section \ref{sec_contribution_pm} & Project management \\ \hline
\end{tabular}
\caption{\label{tab_reference_rq3}Sections and tables with answers to RQ3}
\end{table}
